Reproducible Research: Course Project 1
=======================================

## Loading and preprocessing the data

1. Load the data

``` {r, echo = TRUE}
activity <- read.csv("activity.csv")
```

2. Process/transform the data

``` {r, echo = TRUE}
library(lubridate)

activity$date <- ymd(activity$date)
```


## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

``` {r, echo = TRUE}
stepsEachDay <- tapply(activity$steps, activity$date, sum , na.rm = TRUE)
```

2. Make a histogram of the total number of steps taken each day

``` {r, echo = TRUE}
hist(stepsEachDay, xlab = "Steps taken per day", ylab = "frequency", col = "green", breaks = 25, main = "Total Number of steps taken each day")
```

3. Mean and Median of the total number of steps taken per day

``` {r, echo = TRUE}
mean(stepsEachDay)
median(stepsEachDay)
```



## What is the average daily activity pattern?

1. time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

``` {r, echo = TRUE}
library(ggplot2)
library(dplyr)

activityInterval <- group_by(activity, interval)
AvgSteps <- summarise(activityInterval, average = mean(steps, na.rm = TRUE))

qplot(AvgSteps$interval,AvgSteps$average, geom = "line", xlab = "Interval", ylab = "average number of steps")
```

2. Which 5-minute interval contains the maximum number of steps?

``` {r, echo = TRUE}
AvgSteps[which.max(AvgSteps$average),1]
```



## Imputing missing values

1.  total number of missing values in the dataset

``` {r, echo = TRUE}
NAValues <- nrow(activity[is.na(activity$steps),])
```

2. Create a new dataset that is equal to the original dataset but with the missing data filled in.

``` {r, echo = TRUE}
library(Hmisc)
activityImputed <- activity
activityImputed$steps <- impute(activity$steps, fun=mean)
```

3. Histogram of the total number of steps taken each day after imputation

``` {r, echo = TRUE}
stepsImputed <- tapply(activityImputed$steps, activityImputed$date, sum)
hist(stepsImputed, xlab = "Steps taken per day", ylab = "frequency", col = "green", breaks = 25, main = "Total Number of steps taken each day")
```

4. Mean and Median total number of steps taken per day after imputation

``` {r, echo = TRUE}
mean(stepsImputed)
median(stepsImputed)
```

It is obvious that final results were remarakabely changed after imputation

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend"

``` {r, echo = TRUE}
activityImputed <- mutate(activityImputed, day = factor(weekdays(activityImputed$date)))
activityImputed <- mutate(activityImputed, weekday = factor((day == "Saturday" | day == "Sunday" ), labels = c("Weekday", "Weekend")))
```


2. panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

``` {r, echo = TRUE}
activityImputedInterval <- group_by(activityImputed, interval, weekday)
AvgImputedSteps <- summarise(activityImputedInterval, average = mean(steps))
qplot(interval,average, data = AvgImputedSteps, geom = "line", xlab = "Interval", ylab = "average number of steps", facets = .~ weekday)
```



